\documentclass[titlepage, 11pt]{report}

\usepackage{array} % provides extra options for tables
\usepackage{booktabs} % for enhancing the presentation of tables
\usepackage{fontspec} % for loading arbitrary fonts
\usepackage{geometry} % for changing margins
\usepackage{graphicx} % for loading the cover image
\usepackage[hidelinks]{hyperref} % for making hyperlinks
\usepackage{minted} % for colour-coding code/data snippets
\usepackage{parskip} % for adding spacing between paragraphs
\usepackage{pstricks} % for loading the cover image
\usepackage{titlesec} % for defining the style of headings
\usepackage{tocloft} % for adjusting the spacing of the table of contents
\usepackage{xcolor} % for creating custom colours

\usepackage[T1]{fontenc}
\usepackage{DejaVuSans}
\renewcommand*\familydefault{\sfdefault} % use the sans-serif variant by default
\setmonofont{JetBrains Mono}


\definecolor{spdxblue}{RGB}{0,65,107}
\definecolor{spdxlightblue}{RGB}{67,152,204}


\geometry{
  paper=a4paper,
  top=2cm,
  bottom=4cm,
  left=2cm,
  right=2cm,
}


\parindent=0pt


\titleformat{\chapter}[display]{
  \fontspec{Trueno}\color{spdxlightblue}\huge\bfseries}{Clause \thechapter}{20pt}{}
\titlespacing*{\chapter} {0pt}{50pt}{40pt}

\titleformat{\section}{
  \fontspec{Trueno}\color{spdxblue}\Large\bfseries}{\thesection}{1em}{}
\titlespacing*{\section} {0pt}{3.5ex plus 1ex minus .2ex}{2.3ex plus .2ex}

\titleformat{\subsection}{
  \fontspec{Trueno}\color{spdxblue}\large\bfseries}{\thesubsection}{1em}{}
\titlespacing*{\subsection} {0pt}{3.25ex plus 1ex minus .2ex}{1.5ex plus .2ex}


\renewcommand*{\cftsecnumwidth}{30pt}
\renewcommand*{\cftsubsecnumwidth}{40pt}
% ensures that there is enough space in the table of contents for
% section numbers with lots of digits


\renewcommand*{\cfttoctitlefont}{\fontspec{Trueno}\huge\bfseries\color{spdxlightblue}}
\renewcommand*{\cftchapfont}{\fontspec{Trueno}\large\bfseries\color{spdxblue}}


\begin{document}

\begin{titlepage}
  \centering
  \vspace*{3cm}

  \def\svgwidth{100pt}
  \scalebox{3}{\input{logo.tex}}

  \par\vspace{3cm}

  {\scshape\Large Software Package Data Exchange\textregistered (SPDX\textregistered)\par}

  \vspace{20pt}

  {\scshape\large Specification Version 2.3.0\par}

  \par
\end{titlepage}

\tableofcontents


\chapter*{The Software Package Data Exchange (SPDX) Specification Version 2.3}
\addcontentsline{toc}{chapter}{Copyright}

Copyright \textcopyright 2010-2022 Linux Foundation and its Contributors. This
work is licensed under the Creative Commons Attribution License 3.0 Unported
(CC-BY-3.0) reproduced in its entirety in Annex J herein. All other rights are
expressly reserved.

This PDF edition typeset in 11pt DejaVu Sans by Sebastian Crane.

With thanks to Adam Cohn, Alan Tse, Alexios Zavras, Andrew Back, Ann Thornton,
Basil Peace, Bill Schineller, Bradlee Edmondson, Bruno Cornec, Ciaran Farrell,
Daniel German, David Edelsohn, David Kemp, David A. Wheeler, Debra McGlade,
Dennis Clark, Dick Brooks, Ed Warnicke, Eran Strod, Eric Thomas, Esteban
Rockett, Gary O'Neall, Guillaume Rousseau, Hassib Khanafer, Henk Birkholz,
Hiroyuki Fukuchi, Jack Manbeck, Jaime Garcia, Jeff Luszcz, Jilayne Lovejoy, John
Ellis, Jonas Oberg, Kamsang Salima, Karen Copenhaver, Kate Stewart, Kevin
Mitchell, Kim Weins, Kirsten Newcomer, Kris Reeves, Liang Cao, Lon Hohberger,
Marc-Etienne Vargenau, Mark Gisi, Marshall Clow, Martin Michlmayr, Martin von
Willebrand, Mark Atwood, Matija Šuklje, Matt Germonprez, Maximilian Huber,
Michael J. Herzog, Michel Ruffin, Nisha Kumar, Norio Kobota, Nuno Brito, Oliver
Fendt, Paul Madick, Peter Williams, Phil Robb, Philip Koltun, Philip Odence,
Phillippe Ombredanne, Pierre Lapointe, Rana Rahal, Robert Martin, Robin Gandhi,
Rose Judge, Sam Ellis, Sameer Ahmed, Scott K Peterson, Scott Lamons, Scott
Sterling, Sean Barnum, Sebastian Crane, Shane Coughlan, Steve Cropper, Steve
Winslow, Stuart Hughes, Thomas F. Incorvia, Thomas Steenbergen, Tom Callaway,
Tom Vidal, Toru Taima, Venkata Krishna, W. Trevor King, William Bartholomew, Yev
Bronshteyn, Yoshiko Ouchi, Yoshiyuki Ito, Yuji Nomura and Zachary McFarland for
their contributions and assistance.


\chapter*{Introduction}
\addcontentsline{toc}{chapter}{Introduction}

Companies and organizations (collectively ``Organizations'') are widely using
and reusing open source and other software packages. Accurate identification of
software is key for many supply chain processes. Vulnerability remediation
starts with knowing the details of which version of software is in use on a
system. Compliance with the associated licenses requires a set of analysis
activities and due diligence that each Organization performs independently,
which may include a manual and/or automated scan of software and identification
of associated licenses followed by manual verification. Software development
teams across the globe use the same open source packages, but little
infrastructure exists to facilitate collaboration on the analysis or share the
results of these analysis activities. As a result, many groups are performing
the same work leading to duplicated efforts and redundant information. With this
document, the SPDX workgroup has created a data exchange format so that
information about software packages and related content may be collected and
shared in a common format with the goal of saving time and improving data
accuracy.


\chapter{Scope}

This Software Package Data Exchange (SPDX) specification defines a standard data
format for communicating the component and metadata information associated with
software packages. An SPDX document can be associated with a set of software
packages, files or snippets and contains information about the software in the
SPDX format described in this specification.


\chapter{Normative references}

The following documents are referred to in the text in such a way that some or
all of their content constitutes requirements of this document. For dated
references, only the edition cited applies. For undated references, the latest
edition of the referenced document (including any amendments) applies.

\newcommand{\clausetworeference}[2]{
\item
  \parbox{\linewidth}{#1, \url{#2}}
}

\begin{itemize}
  \clausetworeference{\textbf{Apache Maven}, Apache Software Foundation}{https://maven.apache.org/}
  \clausetworeference{\textbf{Bower API}}{https://bower.io/docs/api/\#install}
  \clausetworeference{\textbf{Common Platform Enumeration (CPE) – Specification}, The MITRE Corporation}{https://cpe.mitre.org/files/cpe-specification_2.2.pdf}
  \clausetworeference{\textbf{Common Platform Enumeration: Naming Specification Version 2.3} (NISTIR 7695), NIST}{https://csrc.nist.gov/publications/detail/nistir/7695/final}
  \clausetworeference{\textbf{npm-package.json}, npm Inc.}{https://docs.npmjs.com/files/package.json}
  \clausetworeference{\textbf{NuGet documentation}, Microsoft}{https://docs.microsoft.com/en-us/nuget/}
  \clausetworeference{\textbf{The Open Group Base Specifications Issue 7} (POSIX.1-2017), 2018 edition, IEEE/Open Group}{https://pubs.opengroup.org/onlinepubs/9699919799/}
  \clausetworeference{\textbf{purl (package URL)}}{https://github.com/package-url/purl-spec}
  \clausetworeference{\textbf{Resource Description Framework (RDF)}, 2014-02-25, W3C}{http://www.w3.org/standards/techs/rdf}
  \clausetworeference{\textbf{The MD5 Message-Digest Algorithm} (RFC-1321), The Internet Society Network Working Group}{https://tools.ietf.org/html/rfc1321}
  \clausetworeference{\textbf{US Secure Hash Algorithm 1 (SHA1)} (RFC-3174), The Internet Society Network Working Group}{https://tools.ietf.org/html/rfc3174}
  \clausetworeference{\textbf{Uniform Resource Identifier (URI): Generic Syntax} (RFC-3986), The Internet Society Network Working Group}{https://tools.ietf.org/html/rfc3986}
  \clausetworeference{\textbf{Augmented BNF for Syntax Specifications: ABNF} (RFC-5234), The Internet Society Network Working Group}{https://tools.ietf.org/html/rfc5234}
  \clausetworeference{\textbf{US Secure Hash Algorithms (SHA and SHA-based HMAC and HKDF)} (RFC-6234), The Internet Society Network Working Group}{https://tools.ietf.org/html/rfc6234}
  \clausetworeference{\textbf{SoftWare Heritage persistent IDentifiers (SWHIDs)}}{https://docs.softwareheritage.org/devel/swh-model/persistent-identifiers.html}
  \clausetworeference{\textbf{SPDX and RDF Ontology}}{http://spdx.org/rdf/ontology/spdx-2-2}
  \clausetworeference{\textbf{SPDX License list}, Linux Foundation}{https://spdx.org/licenses/}
  \clausetworeference{\textbf{SPDX License Exceptions list}, Linux Foundation}{https://spdx.org/licenses/exceptions-index.html}
\end{itemize}


\chapter{Terms and definitions}

For the purposes of this document, the following terms and definitions apply.

ISO and IEC maintain terminological databases for use in standardization at the
following addresses:

\begin{itemize}
\item
  ISO Online browsing platform: available at: \url{https://www.iso.org/obp}
\item
  IEC Electropedia: available at \url{http://www.electropedia.org/}
\end{itemize}

\section*{annotations information section}
\textit{section} (3.9) type, an instance of which contains comments about an
SPDX document, SPDX file, SPDX package, or SPDX snippet

\section*{field}
a piece of information contained in a \textit{section} (3.9)

\section*{file information section}
\textit{section} (3.9) type, an instance of which contains facts specific to
files

\section*{other licensing information detected section}
\textit{section} (3.9) type, an instance of which contains a way to capture
information about and refer to licenses that are not on the SPDX license List

\section*{package}
any unit of content that can be associated with a distribution of software

\section*{package information section}
\textit{section} (3.9) type, an instance of which contains facts that are common
properties of a \textit{package}

\section*{relationships between SPDX elements information section}
\textit{section} (3.9) type, an instance of which contains information on how
documents, \textit{packages} (3.5), files and snippets relate to each other

\section*{review information section}
\textit{section} (3.9) type, an instance of which contains information about
persons, organizations or tools that have reviewed a document

\section*{section}
a part of this SPDX specification

\section*{snippet information section}
\textit{section} (3.9) type, an instance of which contains facts that are
specific to a part of a file

\section*{SPDX document}
collection of \textit{section} (3.8) instances each of which contains
information about software organized using the \textit{SPDX format} (3.11)

\section*{SPDX document creation information section}
\textit{section} (3.9) type, an instance of which contains metadata that
associates analysis results with a specific version of an SPDX document (3.11)
and license for use, and provides information on how, when, and by whom the SPDX
document was created

\section*{SPDX format}
the data format defined by this document

\section*{sub-package}
a package which is embedded in a larger package


\chapter{Conformance}

\section{SPDX Current and Previous Versions}
This edition has the version number 2.3 as part of its title. This is a follow
on from ISO/IEC 5962:2021 Information technology — SPDX\textregistered
Specification V2.2.1, and includes new fields. Earlier editions were published
by the SPDX workgroup via the Linux Foundation. The SPDX Specification was
subsequently transposed into the Joint Development Foundation. (Those earlier
editions are: 1.0 (August 2011), 1.1 (August 2012), 1.2 (October 2013), 2.0 (May
2015), 2.1 (November 2016), and 2.2 (May 2020).) Differences between this
edition and earlier ones are reported in Annex J; see also the Bibliography.

\section{Obsolete features}
Over the life of a standard, some older approaches can become obsolete and are
dropped from subsequent editions, possibly with a replacement approach being
provided. Such action involves \textit{deprecating} those outdated features.
This edition identifies all currently deprecated features. No features present
in 2.2.1 have been deprecated in this 2.3 release.

\section{Alternate notation for some conformance requirements}
This standard contains more than a few \textit{cardinality assertions}, each of
which indicates absolute, optional, or conditional requirements. Here are some
examples:

\begin{itemize}
\item
  Cardinality: Mandatory, one.
\item
  Cardinality: Optional, one or many.
\item
  Cardinality: Mandatory, one if {condition} is true or {feature} omitted, zero
  (shall be omitted) if {condition} is false.
\item
  Cardinality: 0..1
\item
  Cardinality: 0..*
\item
  Cardinality: 1..1
\item
  Cardinality: 1..*
\end{itemize}

Each of these assertions can easily be understood as to whether a feature is
required, and if so, how many occurrences are required; also, whether a feature
is permitted, and if so, in what number. As this is the format long familiar to
the SPDX community, it has been preserved in this document.

\section{Standard data format requirements}
The data format specification and recommendations are subject to the following
constraints:

\begin{itemize}
\item
  Shall be in a human readable form.
\item
  Shall be in a syntax that a software tool can read and write.
\item
  Shall be suitable to be checked for syntactic correctness automatically,
  independent of how it was generated (human or tool).
\item
  The SPDX document character set shall support UTF-8 encoding.
\item
  Multiple serialization formats may be used to represent the information being
  exchanged. Current supported formats include:

  \begin{itemize}
  \item
    \textbf{YAML 1.2} see: \url{https://yaml.org/spec/1.2/spec.html}
    \item
      \textbf{JavaScript Object Notation} (JSON) see:
      \url{https://www.ecma-international.org/publications/files/ECMA-ST/ECMA-404.pdf}
      \begin{itemize}
      \item
        The JSON Schema for SPDX can be found in the SPDX Spec Git Repository
        Schema directory
        (\url{https://github.com/spdx/spdx-spec/blob/master/schemas/spdx-schema.json})
      \end{itemize}
    \item
      \textbf{Resource Description Framework} (RDF also referred to as RDF/XML)
      see: \url{https://www.w3.org/TR/rdf-syntax-grammar/}
    \item
      \textbf{tag:value} flat text file as described in this specification
    \item
      \textbf{.xls} spreadsheets
  \end{itemize}

\item
  In addition to the supported formats, the following format is in development
  with a plan to complete the specification in the next release:
  \begin{itemize}
  \item
    \textbf{Extensible Markup Language} (XML) see:
    \url{https://www.w3.org/TR/2008/REC-xml-20081126/}
  \end{itemize}

\item
  Interoperability between all the supported file formats shall be preserved.
\item
  SPDX defines how to validate a document in each supported format, and how to
  translate a valid document without loss to each other supported format.
\item
  Tags and format properties are case sensitive.
\newpage %% force text item and table onto new page since there isn't enough space.
\item
  Should be easy to recognize in a file system without opening the file. A
  suggested naming convention is:

  \begin{table}[htbp]
    \begin{tabular*}{\linewidth}{>{\itshape}lp{0.75\linewidth}}
      \toprule
      \textup{Format} & \textup{Extension} \\
      \midrule
      tag:value & *.spdx \\
      RDF & *.spdx.rdf \\
      JSON & *.spdx.json \\
      XML & *.spdx.xml \\
      YAML & *.spdx.yaml or *.spdx.yml \\
      \bottomrule
    \end{tabular*}
  \end{table}

\item
  The convention in this specification is for the RDF examples to use
  \texttt{rdf:about="..."} to represent that a proper Uniform Resource Indicator
  (URI) should be present.
\end{itemize}

\section{Trademark Compliance}
To be designated an SPDX document, a file shall comply with the requirements of
the SPDX Trademark License (See the \href{https://spdx.dev/trademark/}{SPDX
Trademark Page}).

The official copyright notice that shall be used with any verbatim reproduction
and/or distribution of this SPDX Specification 2.3 is:

\begin{quote}
  ``Official SPDX Specification 2.3 Copyright \textcopyright 2010-2022 Linux
  Foundation and its Contributors. Licensed under the Creative Commons
  Attribution License 3.0 Unported. All other rights are expressly reserved.''
\end{quote}

The official copyright notice that shall be used with any non-verbatim
reproduction and/or distribution of this SPDX Specification 2.3, including
without limitation any partial use or combining this SPDX Specification with
another work, is:

\begin{quote}
  ``This is not an official SPDX Specification. Portions herein have been
  reproduced from SPDX Specification 2.3 found at spdx.dev. These portions are
  Copyright \textcopyright 2010-2022 Linux Foundation and its Contributors, and
  are licensed under the Creative Commons Attribution License 3.0 Unported by
  the Linux Foundation and its Contributors. All other rights are expressly
  reserved by Linux Foundation and its Contributors.''
\end{quote}

\section{The SPDX Lite profile}
Rather than conforming to this whole specification, an implementation may
conform with SPDX Lite only, a profile that defines a subset of the SPDX
specification. SPDX Lite aims at the balance between the SPDX standard and
actual workflows in some industries. See Annex G for more information.


\chapter{Composition of an SPDX document}

\section{What this specification covers}
This document contains the specification for an SPDX document, which is made up
of a set of one or more sections, instances of which contain information in the
form of \textit{fields}. The following subclauses introduce the different kinds
of sections allowed. The fields for each kind of section are defined in the
clause corresponding to that section. Within an SPDX document, sections may be
organized, as follows:

\textbf{SPDX v2.3 Document shall contain:}
\begin{itemize}
\item
  SPDX Document Creation Information
\end{itemize}

\textbf{SPDX v2.3 Document may contain:}
\begin{itemize}
\item
  Package Information
\item
  File information
\item
  Snippet Information
\item
  Other Licensing Information Detected
\item
  Relationships between SPDX Elements information
\item
  Annotations Information
\item
  Review information
\end{itemize}

The object model is illustrated by Annex C.

\section{Sections}

\subsection{SPDX document creation information section}
An instance of this section provides the necessary information for forward and
backward compatibility for processing tools.

One instance shall be present for each SPDX document produced.

Cardinality: Mandatory, one.

See Clause 6 for details of the fields in this kind of section.

\subsection{Package information section}
If SPDX information is being used to describe packages, then one instance of the
package information per package being described shall exist. It provides
important meta information about the package as a whole. Packages are an
abstract concept that can be used to refer to any distribution of software,
typically consisting of one or more files and capable of containing
sub-packages. Starting with SPDX 2.0, it is not necessary to have a package
wrapping a set of files.

A package refers to any unit of content that can be associated with a
distribution of software. Typically, a package is composed of one or more
files. An SPDX document may, but is not required to, provide details about the
individual files comprising a package (see Clause 8).

Any of the following non-limiting examples may be (but are not required to be)
represented in SPDX as a package:

\begin{itemize}
\item
  a tarball, zip file or other archive
\item
  a directory or sub-directory
\item
  a separately distributed piece of software which another Package or File uses
  or depends upon (e.g., a Python package, a Go module, …)
\item
  a container image, and/or each image layer within a container image
\item
  a collection of one or more sub-packages
\item
  a Git repository snapshot from a particular point in time
\end{itemize}

Note that some of these could be represented in SPDX as a file as well.

In an SPDX document, relationship elements can be used to indicate relationships
between packages, such as dependency relationships.

Cardinality: Optional, zero or many.

See Clause 7 for details of the fields in this kind of section.

In \texttt{tag:value} format, the order in which package and files occur is
syntactically significant.

\begin{itemize}
\item
  A new package Information section is denoted by the package name (7.1) field.
\item
  All package information fields shall be grouped together before the start of a
  files section (Clause 8), if file(s) are present.
\item
  All files contained in a package shall immediately follow the applicable
  package information.
\item
  A new package information section (via package name) denotes the start of
  another package.
\item
  Sub-packages shall not be nested inside a package information section, but
  shall be separate and shall use a relationship to clarify.
\item
  Annotations and relationships for the package may appear after the package
  information before any file information.
\end{itemize}

\subsection{File information section}
One instance of the file information may exist for each file in the software
package. It provides important meta information about a given file including
licenses and copyright. Starting with SPDX 2.0, it is not necessary to have a
package wrapping a set of files.

Cardinality: Optional, zero or many.

See Clause 8 for details of the fields in this kind of section.

When implementing \texttt{tag:value} format, the positioning of file elements is
syntactically significant:

\begin{itemize}
\item
  Files are assumed to be associated with the package information that
  immediately precedes it, if a package exists
\item
  Presence of a new package information signals the end of the set of files
  associated with the preceding package, unless an explicit relationship is used
\item
  If a package contains files, the file information sections shall follow its
  package information section
\item
  If a file is not part of any package, it shall precede any package information
  section reference in the SPDX document
\item
  The first field to start off the description of a file shall be the file name
  in \texttt{tag:value} format
\item
  File information is associated with the file name that precedes it
\item
  Annotations on the file and relationships from the file may appear after the
  file information, before the next file or package information section.
\end{itemize}

When implementing file information in RDF, the \texttt{spdx:hasFile} property is
used to associate the package with the file.

\subsection{Snippet information section}
Snippets can optionally be used when a file is known to have some content that
has been included from another original source. They are useful for denoting
when part of a file may have been originally created under another license or
copied from a place with a known vulnerability.

Each instance of Snippet Information shall be associated with a specific file in
an SPDX document.

Cardinality: Optional, zero or many.

See Clause 9 for details of the fields in this kind of section.

When implementing \texttt{tag:value} format, the positioning of snippet elements
is syntactically significant:

\begin{itemize}
\item
  If a file contains snippets, the snippet information section shall follow a
  related file information section (if it exists in the document).
\item
  Presence of a new file or package section signals the end of the set of
  snippets associated with the original file, unless an explicit relationship is
  used.
\item
  The first field to start off the description of a snippet shall be the snippet
  identifier in \texttt{tag:value} format.
\item
  Annotations on the snippet and relationships from the snippet may appear after
  the snippet information, before the next file or package section.
\end{itemize}

\subsection{Other licensing information detected section}
This section is used for any detected, declared or concluded licenses that are
NOT on the SPDX License List. For the most up-to-date version of the list, see
\url{https://spdx.org/licenses/}. The SPDX License List can also be found in
Annex A.

One instance shall be created for every unique license or licensing information
being referenced that does not match one of the licenses on the SPDX License
List.

Cardinality: Optional, zero or many.

See Clause 10 for details of the fields in this kind of section.

\subsection{Relationships between SPDX elements information section}
Packages, files, and snippets are all considered to be SPDX elements, and
relationships can be made explicit between these SPDX elements by using the
fields in this section.

Cardinality: Optional, zero or many.

See Clause 11 for details of the fields in this kind of section.

\subsection{Annotations information section}
Annotations permit the addition of information to validate and clarify ambiguous
SPDX elements (packages, files or snippets).

Cardinality: Optional, zero or many.

See Clause 12 for details of the fields in this kind of section.

This section is now the preferred home for review information.

\subsection{Review information section}
The review information section is included for compatibility with SPDX 1.2, and
is deprecated since SPDX 2.0. Any review information shall use an annotation (as
described in Clause 12) with an annotation type of \texttt{REVIEW}.

Review information may be added after the initial SPDX document has been
created. The set of fields are optional and multiple instances may be
added. Once a reviewer entry is added, the review date associated with the
review is mandatory. The created date shall not be modified as a result of the
addition of information regarding the conduct of a review. A review comments is
optional.

See Clause 13 for details of the fields in this kind of section.

\section{What this specification does not cover}
This document does not address the following:

\begin{itemize}
\item
  Information that cannot be derived from an inspection (whether manual or using
  automated tools) of the package to be analyzed.
\item
  How the data stored in an SPDX document is used by the recipient.
\item
  Any identification of any patent(s) which may or may not relate to the
  package.
\item
  Legal interpretation of the licenses or any compliance actions that have been
  or may need to be taken.
\end{itemize}


\end{document}
